---
title: "Luke's Version"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
```

\## \*\*Using iNaturalist to Identify Breeding Locations of Manta Ray\*\*

Authors: Luke Engstrom

Co-Author: Dr. Christopher Merkord

Affiliations: Minnesota State University of Moorhead

\## Abstract

\- Manta ray migrate where the food is, which makes them hard to study​

\- In addition, they are at risk from fisheries and bycatch (Stewart et al 2016)​

\- Using iNaturalist data to see if there are migration patterns​

\- Then identifying breeding grounds of two different species of manta ray​

\- With this information we propose to make these sites protected grounds

\## Introduction

\- Manta ray are an elusive species, juveniles have never been studied in the wild (Stewart et al 2018)​

\- Mobula birostris (oceanic ray) and Mobula alfredi (reef ray) are both migratory species​

\- Reef Rays typically have their mating season from October to January (Simpkins 2013)​

\- Oceanic Rays are considered year-round, but it depends on what waters they are in (Manta Ray Reproduction 2018)​

\- We are working to identify breeding grounds to propose marine protected areas

\## Methods

\- Using iNaturalist data, a citizen science program used to catalog sightings of organisms​

\- Then cleaned up the data using rNat and tidyverse​

\- We filtered the data to include species name, latitude, longitude, year, month, and day​

\- Then made a map with the filtered data​

\## Results

\- Oceanic rays may have breeding grounds near the west coast of the US, Gulf of Mexico, New Zealand, Australia, and Indonesia​

\- Reef rays may have breeding grounds near India in the Maldives and Indonesia​

\- Reef rays have higher sightings in winter near India in the Maldives and Indonesia​

```{r cars}
summary(cars)
```

```{r pressure, echo=FALSE}
plot(pressure)
```
